---
output: html_document
---
#E affinis 16S analysis
### Martin Bontrager

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs_abundance/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r read_data}
library(phyloseq)
library(RColorBrewer)
library(ggplot2)
library(ape)
library(plyr)

set.seed(3399)

otuTable <- read.csv("data/otus.csv", header = TRUE, row.names = 1)
taxTable <- read.csv("data/taxa.csv", quote = '"')
sampleData <- read.csv("data/sampleData.csv")
tree <- read.tree("data/tree.tre")
rownames(sampleData) <- sampleData[, 1]
sampleData <- sampleData[, -1]
taxmat <- as.matrix(taxTable)
rownames(taxmat) <- taxmat[, 1]
taxmat <- taxmat[, -1]


otuTable <- otu_table(otuTable, taxa_are_rows = FALSE)
taxTable <- tax_table(taxmat)

physeq <- phyloseq(otuTable, taxTable, tree)
sample_data(physeq) <- sampleData

## Standardize abundances by relative abundance for visualizations:
physeq_relative <- transform_sample_counts(physeq, function(x) x / sum(x))

rm(taxmat, sampleData, otuTable, taxTable, tree)

water <- c("BBW", "BRW", "CBW", "EUW", "FRW", "IJW", "LOW", "MAW", "MMW", "POW", "RMW", "SCW", "TXW", "VIW")
copepod <- c("BRE", "BBE", "CBE", "EUE", "FRE", "IJE", "LOE", "MAE", "MME", "POE", "RME", "SCE", "TXE", "VIE")

# ## ggplot theming
theme_set(theme_bw())
pal = "Set1"
scale_colour_discrete <- function(palname = pal, ...) {
        scale_colour_brewer(palette = palname, ...)
}

scale_fill_discrete <- function(palname = pal, ...) {
        scale_fill_brewer(palette = palname, ...)
}

getPalette <- colorRampPalette(brewer.pal(9, "Set1"))

# Plot title size, etc.
fontsize = 18L
theme_update(axis.title.x = element_text(size=fontsize))
theme_update(axis.title.y = element_text(size=fontsize))
theme_update(plot.title = element_text(size=fontsize+2))
```

#Abundance Bar Plots


I have transformed the data by relative abundance. I am plotting only the most abundance phyla and the composition of the phyla.

## Abundance by Phylum

``` {r abundance_phylum}
theme_set(theme_bw())

# Plotting function. Feed in a phyloseq object and a color value object
comparison_plot <- function(matr, color_val, label = "Sample", fill = "Phylum", title = NULL){
    aa <- psmelt(matr)
    .e <- environment()
    
    p <- ggplot(aa, aes_string(x = label, y = "Abundance", fill = fill), environment = .e) +
    geom_bar(aes(order=desc(aa[,fill])),
                     stat = "identity", 
                     position = "stack", 
                     color = "black") +
    scale_fill_manual(values = color_val) +
    guides(fill=guide_legend(ncol=2)) +
    theme(axis.title.y = element_text(angle=90, size=20),
          axis.title.x = element_text(size=20),
          axis.text.x = element_text(angle = 90, hjust=0, vjust=0.5, size=18),
          plot.title = element_blank())
    
    return(p)
}

# I only look at the top 10 phyla
physeq_glom <- tax_glom(physeq, "Phylum")
physeq_rel_glom <- transform_sample_counts(physeq_glom, function(x) x / sum(x))
a <- prune_taxa(names(sort(taxa_sums(physeq_rel_glom), decreasing = TRUE)[1:10]), 
                physeq_rel_glom)
colorCount <- length(levels(as.data.frame(tax_table(a))$Phylum))
color_val <- sample(getPalette(colorCount))

comparison_plot(a, color_val, label = "Code", title = "Abundance by Phylum") +
    guides(fill=guide_legend(ncol=1)) + 
    ggtitle(title)

cop <- subset_samples(a, Environment=="Copepod")
wat <- subset_samples(a, Environment=="Water")

#Plotting (to get the bar order consistent)
comparison_plot(cop, color_val, label = "Code", title = "Abundance by Phylum in Copepods") +
    guides(fill=guide_legend(ncol=1)) + 
    ggtitle(title)
comparison_plot(wat, color_val, label = "Code", title = "Abundance by Phylum in Water") +
    guides(fill=guide_legend(ncol=1)) + 
    ggtitle(title)
```

#Test for unequal variances at the phylum level

```{r unequal_var_phylum}
library(reshape2)
#Get top 10 phyla for each water and copepods
cop <- subset_samples(physeq_rel_glom, Environment=="Copepod")
cop <- prune_taxa(names(sort(taxa_sums(cop), decreasing = TRUE)[1:10]), cop)
wat <- subset_samples(physeq_rel_glom, Environment=="Water")
wat <- prune_taxa(names(sort(taxa_sums(wat), decreasing = TRUE)[1:10]), wat)

a <- psmelt(wat)[, 1:3]
b <- psmelt(cop)[, 1:3]
ab <- rbind(a, b)
c <- dcast(ab, OTU ~ Sample, value.var = "Abundance")
e <- as.data.frame(tax_table(physeq))
e <- cbind(OTU = rownames(e), e)
f <- merge(e, c, by = "OTU")
d <- sapply(f[,1], function(x) gsub("OTU_", "", x), USE.NAMES = FALSE)
d <- as.numeric(d)
f$OTU <- sprintf("OTU_%05d", d)
f <- f[order(f$OTU),]
wrong_cols <- c("OTU", "Kingdom", "Class", "Order", "Family", "Genus", "Species")
f <- f[, !(colnames(f) %in% wrong_cols)]
rownames(f) <- f[,1]
f$Phylum <- NULL
f <- t(f)
write.csv(f, "output/sample_matrix_phylum_relative.csv")

c <- f[(rownames(f) %in% copepod), ]
c <- c[, colSums(is.na(c)) != nrow(c)]
c <- c[, order(colSums(c), decreasing = TRUE)]
w <- f[(rownames(f) %in% water), ]
w <- w[, colSums(is.na(w)) != nrow(w)]
w <- w[, order(colSums(w), decreasing = TRUE)]

for(i in 1:10){
    print(var.test(c[,i], w[,i]))
}

```

#Abundance by phylum in merged samples

``` {r merged, fig.width=6, fig.height=4}
# Bar plots of abundance for merged copepod vs. water samples
merged_env <- merge_samples(physeq, "Environment")
merged_env <- transform_sample_counts(merged_env, function(x) x / sum(x))
merged_env_rel <- tax_glom(merged_env, "Phylum")
a <- prune_taxa(names(sort(taxa_sums(merged_env_rel), decreasing = TRUE)[1:10]), 
                merged_env_rel)
pdf("fig-phylum_abundance_by_env.pdf", width=6, height=4)
comparison_plot(a, color_val) + 
    guides(fill=guide_legend(ncol=1)) + 
     theme(axis.title.y = element_text(angle=90),
          axis.title.x = element_blank(),
          axis.text.x = element_text(angle=0, hjust=0.5, vjust=0.5),
          plot.title = element_blank())
dev.off()
```

## Abundance by Class

```{r abundance_class}
# I only look at the top 20 Classes
physeq_glom <- tax_glom(physeq, "Class")
physeq_rel_glom <- transform_sample_counts(physeq_glom, function(x) x / sum(x))
a <- prune_taxa(names(sort(taxa_sums(physeq_rel_glom), decreasing = TRUE)[1:10]), 
                physeq_rel_glom)
colorCount <- length(levels(as.data.frame(tax_table(a))$Class))
color_val <- sample(getPalette(colorCount))

# Subset samples for plotting
cop <- subset_samples(a, Environment=="Copepod")
wat <- subset_samples(a, Environment=="Water")

#Plotting (to get the bar order consistent)
pdf("figS-Abundance_class_copepods.pdf", height=6, width=9)
comparison_plot(cop, color_val, label = "Code", fill = "Class") +
    guides(fill=guide_legend(ncol=1)) 
dev.off()

pdf("figS-Abundance_class_water.pdf", height=6, width=9)
comparison_plot(wat, color_val, label = "Code", fill = "Class") +
    guides(fill=guide_legend(ncol=1))
dev.off()
```

#Test unequal variance in Classes

```{r unequal_var_class}
library(reshape2)

#Get top 10 phyla for each water and copepods
cop <- subset_samples(physeq_rel_glom, Environment=="Copepod")
cop <- prune_taxa(names(sort(taxa_sums(cop), decreasing = TRUE)[1:10]), cop)
wat <- subset_samples(physeq_rel_glom, Environment=="Water")
wat <- prune_taxa(names(sort(taxa_sums(wat), decreasing = TRUE)[1:10]), wat)

a <- psmelt(wat)[, 1:3]
b <- psmelt(cop)[, 1:3]
ab <- rbind(a, b)
c <- dcast(ab, OTU ~ Sample, value.var = "Abundance")
e <- as.data.frame(tax_table(physeq))
e <- cbind(OTU = rownames(e), e)
f <- merge(e, c, by = "OTU")
d <- sapply(f[,1], function(x) gsub("OTU_", "", x), USE.NAMES = FALSE)
d <- as.numeric(d)
f$OTU <- sprintf("OTU_%05d", d)
f <- f[order(f$OTU),]
wrong_cols <- c("OTU", "Kingdom", "Phylum", "Order", "Family", "Genus", "Species")
f <- f[, !(colnames(f) %in% wrong_cols)]
f$Class <- as.character(f$Class)
f$Class[4] <- "unclassified_1"
f$Class[10:11] <- c("unclassified_2", "unclassified_3")
rownames(f) <- f[,1]
f$Class <- NULL
f <- t(f)
write.csv(f, "output/sample_matrix_class_relative.csv")

c <- f[(rownames(f) %in% copepod), ]
c <- c[, colSums(is.na(c)) != nrow(c)]
c <- c[, order(colSums(c), decreasing = TRUE)]
w <- f[(rownames(f) %in% water), ]
w <- w[, colSums(is.na(w)) != nrow(w)]
w <- w[, order(colSums(w), decreasing = TRUE)]

for(i in 1:10){
    print(var.test(c[,i], w[,i], alternative = "greater"))
}

library(matrixStats)
a <- colVars(c)
b <- colVars(w)

t.test(a, b, alternative="greater")
```

## Abundance by Order

```{r abundance_order}
# I only look at the top 32 taxa
physeq_glom <- tax_glom(physeq, "Order")
physeq_rel_glom <- transform_sample_counts(physeq_glom, function(x) x / sum(x))
a <- prune_taxa(names(sort(taxa_sums(physeq_rel_glom), decreasing = TRUE)[1:32]), 
                physeq_rel_glom)
colorCount <- length(levels(as.data.frame(tax_table(a))$Order))
color_val <- sample(getPalette(colorCount))

# Subset samples for plotting
cop <- subset_samples(a, Environment=="Copepod")
wat <- subset_samples(a, Environment=="Water")

#Plotting (to get the bar order consistent)
comparison_plot(cop, color_val, label = "Code", 
                title = "Abundance by Order in Copepods", fill = "Order") + 
    ggtitle(title)
comparison_plot(wat, color_val, label = "Code", title = "Abundance by Order in Water", 
                fill = "Order") + 
    ggtitle(title)
```

## Abundance by Family

```{r abundance_family}
# I only look at the top 32 taxa
physeq_glom <- tax_glom(physeq, "Family")
physeq_rel_glom <- transform_sample_counts(physeq_glom, function(x) x / sum(x))
a <- prune_taxa(names(sort(taxa_sums(physeq_rel_glom), decreasing = TRUE)[1:32]), 
                physeq_rel_glom)
colorCount <- length(levels(as.data.frame(tax_table(a))$Family))
color_val <- sample(getPalette(colorCount))

# Subset samples for plotting
cop <- subset_samples(a, Environment=="Copepod")
wat <- subset_samples(a, Environment=="Water")

#Plotting (to get the bar order consistent)
comparison_plot(cop, color_val, label = "Code", title = "Abundance by Family in Copepods", 
                fill = "Family") + 
    ggtitle(title)
comparison_plot(wat, color_val, label = "Code", title = "Abundance by Family in Water", 
                fill = "Family") + 
    ggtitle(title)
```

## Abundance by genus

```{r abundance_genus}
# I only look at the top 32 taxa
physeq_glom <- tax_glom(physeq, "Genus")
physeq_rel_glom <- transform_sample_counts(physeq_glom, function(x) x / sum(x))
a <- prune_taxa(names(sort(taxa_sums(physeq_rel_glom), decreasing = TRUE)[1:32]), 
                physeq_rel_glom)
colorCount <- length(levels(as.data.frame(tax_table(a))$Genus))
color_val <- sample(getPalette(colorCount))

# Subset samples for plotting
cop <- subset_samples(a, Environment=="Copepod")
wat <- subset_samples(a, Environment=="Water")

#Plotting (to get the bar order consistent)
comparison_plot(cop, color_val, label = "Code", title = "Abundance by Genus in Copepods", 
                fill = "Genus") + 
    ggtitle(title)
comparison_plot(wat, color_val, label = "Code", title = "Abundance by Genus in Water", 
                fill = "Genus") + 
    ggtitle(title)
```

## Without common phlya

Now I think it might be interesting to look at these plots excluding all actinobacteria, proteobacteria, and bacteroides (The vast majority of taxa belong to those phyla)

```{r remove_common}
theme_set(theme_bw())
physeq_glom <- tax_glom(physeq, "Phylum")
physeq_rel_phyla <- transform_sample_counts(physeq_glom, function(x) x / sum(x))
physeq_uncommon_phyla <- subset_taxa(physeq_rel_phyla, 
                        (Phylum != "Proteobacteria" & 
                         Phylum != "Actinobacteria" & 
                         Phylum != "Bacteroidetes"))

colorCount <- length(levels(as.data.frame(tax_table(physeq_uncommon_phyla))$Phylum))
color_val <- sample(getPalette(colorCount))

cop <- subset_samples(physeq_uncommon_phyla, Environment=="Copepod")
wat <- subset_samples(physeq_uncommon_phyla, Environment=="Water")

comparison_plot(cop, color_val, label = "Code")
comparison_plot(wat, color_val, label = "Code")
```


##Abundance within common phyla

```{r proteobacteria}
theme_set(theme_bw())
proteos <- subset_taxa(physeq_relative, Phylum=="Proteobacteria")
proteos_by_class <- tax_glom(proteos, "Class")
colorCount <- length(levels(as.data.frame(tax_table(proteos_by_class))$Class))
color_val <- sample(brewer.pal(colorCount, "Set1"))

cop <- subset_samples(proteos_by_class, Environment=="Copepod")
wat <- subset_samples(proteos_by_class, Environment=="Water")

#Comparison plots
comparison_plot(cop, color_val, label = "Code", 
                title = "Proteobacteria Abundance in Copepods", 
                fill = "Class")

title <- "Proteobacteria Abundance in Water"
comparison_plot(wat, color_val, label = "Code", title = title, fill = "Class")

proteos <- subset_taxa(physeq_relative, Phylum=="Proteobacteria")
proteos_by_family <- tax_glom(proteos, "Family")
colorCount <- length(levels(as.data.frame(tax_table(proteos_by_family))$Family))
color_val <- sample(getPalette(colorCount))

cop <- subset_samples(proteos_by_family, Environment=="Copepod")
wat <- subset_samples(proteos_by_family, Environment=="Water")

title <- "Proteobacteria Abundance in Copepods by Family"
comparison_plot(cop, color_val, label = "Code", title = title, fill = "Family")

title <- "Proteobacteria Abundance in Water by Family"
comparison_plot(wat, color_val, label = "Code", title = title, fill = "Family")

```

```{r actinomycetes}
# Do the same for actinomycetes
actinos <- subset_taxa(physeq_relative, Phylum=="Actinobacteria")
actinos_by_class <- tax_glom(actinos, "Class")
a <- as.vector(tax_table(actinos_by_class)[, "Class"])
colorCount <- length(unique(a))
color_val <- sample(brewer.pal(colorCount, "Set1"))

cop <- subset_samples(actinos_by_class, Environment=="Copepod")
wat <- subset_samples(actinos_by_class, Environment=="Water")

title <- "Actinobacteria Abundance in Copepods"
comparison_plot(cop, color_val, label = "Code", title = title, fill = "Class")

title <- "Actinobacteria Abundance in Water"
comparison_plot(wat, color_val, label = "Code", title = title, fill = "Class")

#Now plot actinos by family
actinos <- subset_taxa(physeq_relative, Phylum=="Actinobacteria")
actinos_by_family <- tax_glom(actinos, "Family")
colorCount <- length(levels(as.data.frame(tax_table(actinos_by_family))$Family))
color_val <- sample(getPalette(colorCount))

cop <- subset_samples(actinos_by_family, Environment=="Copepod")
wat <- subset_samples(actinos_by_family, Environment=="Water")

title <- "Actinobacteria Abundance in Copepods"
comparison_plot(cop, color_val, label = "Code", title = title, fill = "Family")

title <- "Actinobacteria Abundance in Water"
comparison_plot(wat, color_val, label = "Code", title = title, fill = "Family")
```


``` {r bacteroidetes}
# Do the same for Bacteroidetes
bact <- subset_taxa(physeq_relative, Phylum=="Bacteroidetes")
bact_by_class <- tax_glom(bact, "Class")
a <- as.vector(tax_table(bact_by_class)[, "Class"])
colorCount <- length(unique(a))
color_val <- sample(getPalette(colorCount))

cop <- subset_samples(bact_by_class, Environment=="Copepod")
wat <- subset_samples(bact_by_class, Environment=="Water")

title <- "Bacteroidetes Abundance in Copepods"
comparison_plot(cop, color_val, label = "Code", title = title, fill = "Class")

title <- "Bacteroidetes Abundance in Water"
comparison_plot(wat, color_val, label = "Code", title = title, fill = "Class")

```
