---
output: html_document
---
#E affinis 16S analysis
### Martin Bontrager

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs_diversity_improve/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r read_data}
library(phyloseq)
library(RColorBrewer)
library(ggplot2)
library(ape)
library(reshape2)
library(dplyr)
library(knitr)
library(gridExtra)
library(grid)

set.seed(9373399)

otuTable <- read.csv("data/otus.csv", header = TRUE, row.names = 1)
taxTable <- read.csv("data/taxa.csv", quote = '"')
sampleData <- read.csv("data/sampleData.csv")
tree <- read.tree("data/tree.tre")
rownames(sampleData) <- sampleData[, 1]
sampleData <- sampleData[, -1]
taxmat <- as.matrix(taxTable)
rownames(taxmat) <- taxmat[, 1]
taxmat <- taxmat[, -1]


otuTable <- otu_table(otuTable, taxa_are_rows = FALSE)
taxTable <- tax_table(taxmat)

physeq <- phyloseq(otuTable, taxTable, tree)
sample_data(physeq) <- sampleData

## Relative Abundance
#physeq <- rarefy_even_depth(physeq)

rm(taxmat, sampleData, otuTable, taxTable, tree)

theme_set(theme_bw())
pal = "Set1"
scale_colour_discrete <- function(palname = pal, ...) {
        scale_colour_brewer(palette = palname, ...)
}

scale_fill_discrete <- function(palname = pal, ...) {
        scale_fill_brewer(palette = palname, ...)
}

getPalette <- colorRampPalette(brewer.pal(9, "Set1"))

get_legend<-function(myggplot){
  tmp <- ggplot_gtable(ggplot_build(myggplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

```

# Alpha Diversity Plots

The first step is to understand within-sample diversity and how it differs between samples. Which samples have the largest diversity? Which have the lowest? Are there patterns? We are unsure of what to expect regarding whether water, gut, or copepod samples have the lowest diversity. Althouogh we do expect copepod gut samples to be equally or less diverse that whole animal samples, since the guts should be a subset of the whole animal.

```{r richness_by_env}
theme_set(theme_bw())
colorCount <- length(levels(sample_data(physeq)$Location))
color_val <- sample(getPalette(colorCount))

# Plotting function:
richness_plot <- function(phy, color_val, xval = "Environment", col = "Location", meas = c("Shannon", "Chao1"), title = "Alpha Diversity"){
    
    p <- plot_richness(phy, x = xval, color = col, measures = meas) + 
        scale_colour_manual(values = color_val) +
        geom_point(size = 1.5, alpha = 1.0) + 
        theme(axis.title.y = element_text(angle=90),
              axis.text.x = element_text(angle=90),
              plot.title = element_blank()) + 
        ylab("Diversity")
    
    return(p)
}

#Plot alpha diversity estimates across samples
alpha1 <- richness_plot(physeq, color_val, meas = c("Observed", "Chao", "Simpson", "Shannon"))

alpha1 + geom_boxplot(data=alpha1$data, 
                      aes(x=Environment, y=value, color=NULL)) + 
    geom_point(size = 2, alpha = 1.0)

color_val <- c("red", "blue")
richness_plot(physeq, color_val, col = "Environment", meas = c("Observed", "Chao1", "Simpson", "Shannon")) + 
    geom_boxplot(data=alpha1$data, aes(x=Environment, y=value, color=NULL), alpha=0.5) +
    geom_point(size = 3.5, alpha = 1.0) + theme(axis.text.x = element_blank())

# Create a table of alpha diversity estimates
adiv <- estimate_richness(physeq, measures=c("Chao1", "Shannon", "Observed", 
                                             "Simpson","ACE"))
# Print a table of alpha diversity estimates
kable(adiv)

#Add metadata to diversity estimate table and remove Gut, unpaired water/cop samples
adiv <- merge(adiv, sample_data(physeq), by="row.names")
adiv_all <- adiv

# Filter out just water and copepod samples
adivW <- filter(adiv, Environment == "Water")
adivC <- filter(adiv, Environment == "Copepod")
```

```{r squish_figure, fig.width=8, fig.height=2.6}
color_val <- brewer.pal(3, "Set1")
pdf("fig-alpha_diversity_env_chao.pdf", width = 8, height =2.6)
richness_plot(physeq, color_val, col = "Environment", meas = c("Observed", "Chao", "Simpson", "Shannon")) + 
    geom_boxplot(data=alpha1$data, aes(x=Environment, y=value, color=NULL), 
                 show.legend=FALSE) +
    geom_point(size = 2) +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_blank())
dev.off()
```

### T test to determine whether diversity is greater in water or in copepods

```{r ttest}
# Test whether water diversity is greater than zero
t.test(adivW$Chao1, adivC$Chao1, alternative="greater")
```

### Paired water/copepod Figure and t-test

```{r pairedT, fig.width=6, fig.height=4}
## Plotting the data
colorCount <- length(levels(sample_data(physeq)$Location))
color_val <- sample(getPalette(colorCount))
pdf("figS-Paired_water_copepod_diversity.pdf", width = 6, height =4)
g <- ggplot(adiv, aes(x = Environment, y = Observed, group = factor(Location)))
g <- g + geom_line(size = 1, aes(colour = Location)) + 
        geom_point(size =2, pch = 21, fill = "black") +
        scale_colour_manual(values = color_val) +
    ylab("Observed OTUs")
g
dev.off()
g

#Paired t-test (paired water/copepod) MUST EXCLUDE MONTMAGNAY AND VE BECAUSE they are unpaired
t.test(adivW$Chao1, adivC$Chao1, paired=TRUE, alternative="greater")
```

# Correlation between water and copepod diversity

```{r corplot, fig.width=6, fig.height=4}
# Plot correlation between water diversity and copepod diversity
a <- melt(adiv, id.vars=c("Location", "Environment"), 
          measure.vars=c("Shannon", "Chao1", "Observed"))
b <- dcast(a, Location ~ variable + Environment)
pdf("fig-Chao_correlation.pdf", width=6, height=4)
div_corplot <- ggplot(b, aes(Chao1_Copepod, Chao1_Water)) + 
    geom_point(size=3) + xlab("Copepod Diversity") +
    ylab("Water Diversity") + theme(axis.title.y = element_text(angle=90)) +
    geom_smooth(method=lm, fill = "lavenderblush3") + 
    theme(axis.title.y = element_text(size=14),
          axis.title.x = element_text(size=14),
          plot.title = element_blank())
div_corplot
dev.off()
div_corplot

with(b, cor.test(Chao1_Copepod, Chao1_Water))
```


##Test water vs. water salinity
Test whether diversity in water is affected by salinity

```{r waterVwater}
adivWsalt <- filter(adivW, WaterType =="Salt")
adivWfresh <- filter(adivW, WaterType == "Fresh")

t.test(adivWsalt$Chao1, adivWfresh$Chao1)
```

## Test copepod diversity by salinity
Test whether diversity in copepods is affected by salinity

```{r copepodVcopepod}
adivCsalt <- filter(adivC, WaterType == "Salt")
adivCfresh <- filter(adivC, WaterType == "Fresh")

t.test(adivCsalt$Chao1, adivCfresh$Chao1)
```

#Alpha Diversity Estimates by Salinity

Plot richness of copepods and water in fresh or salt water. Are there patterns of richness depending on salinity? 

```{r richness_by_habitat, fig.width=8, fig.height=2.6}
theme_set(theme_bw())

# Subset data for plotting purposes
copepods_only <- subset_samples(physeq, Environment=="Copepod")
water_only <- subset_samples(physeq, Environment=="Water")

t <- "Copepod Associated Samples"
color_val <- c("lightpink2", "darkred")
rplot_cop <- richness_plot(copepods_only, color_val, xval = "WaterType", 
                           col = "WaterType", title = t, meas = c("Shannon"))
p1 <- rplot_cop + geom_boxplot(data=rplot_cop$data, aes(x=WaterType, y=value, color=NULL), show.legend=FALSE) + 
    geom_point(size = 3) +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          strip.text.x = element_text(size = 12)) +
    ylim(1.5, 5.5)
p1

t <- "Water bacteria samples"
color_val <- c("lightblue", "darkblue")
rplot_water <- richness_plot(water_only, color_val,xval = "WaterType", 
                             col = "WaterType", title = t, meas = c("Shannon")) 
p2 <- rplot_water + geom_boxplot(data=rplot_water$data, aes(x=WaterType, y=value, color=NULL), show.legend=FALSE) +
    geom_point(size=3) +
    theme(axis.title.y = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          strip.text.x = element_text(size = 12),
          legend.text = element_text(size=10)) +
    ylim(1.5, 5.5)

p1 <- p1 + theme(legend.position="none")
p2 <- p2 + theme(legend.position="none")

grid.arrange(p1, p2, ncol=2)

pdf("fig-alpha_diversity_Salinity.pdf", width=4, height=2.6)
grid.arrange(p1, p2, ncol=2)
dev.off()
```

#Alpha Diversity Estimates by Clade

```{r richness_by_clade}
theme_set(theme_bw())
color_val <- c("purple3", "red", "forestgreen")

t <- "Copepod Associated Samples"
rplot_cop <- richness_plot(copepods_only, color_val, xval = "Clade", 
                           col = "Clade", title = t)
p1 <- rplot_cop + theme(axis.title.y = element_text(size=12),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          strip.text.x = element_text(size = 12),
          legend.text = element_text(size=10))
#          plot.title = element_text(size=18, face="bold", vjust=2)) + 
#    ggtitle("Copepod Samples")

t <- "Water bacteria samples"
rplot_water <- richness_plot(water_only, color_val,xval = "Clade", 
                             col = "Clade", title = t)
p2 <- rplot_water + theme(axis.title.y = element_text(size=12),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          strip.text.x = element_text(size = 12),
          legend.text = element_text(size=10))
#          plot.title = element_text(size=18, face="bold", vjust=2)) + 
#    ggtitle("Water Samples")

```

#Alpha Diversity Estimates by Continent

```{r richness_by_continent, fig.width=8, fig.height=5}
theme_set(theme_bw())
color_val1 <- c("dodgerblue3", "goldenrod3")

t <- "Copepod Associated Samples"
rplot_cop <- richness_plot(copepods_only, color_val1, xval = "Continent", 
                           col = "Continent", title = t)
p3 <- rplot_cop + theme(axis.title.y = element_text(size=12),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          strip.text.x = element_text(size = 12),
          legend.text = element_text(size=10))
#          plot.title = element_text(size=18, face="bold", vjust=2)) + 
#    ggtitle("Copepod Samples")

t <- "Water bacteria samples"
rplot_water <- richness_plot(water_only, color_val1,xval = "Continent", 
                             col = "Continent", title = t)
p4 <- rplot_water + theme(axis.title.y = element_text(size=12),
          axis.title.x = element_blank(),
          axis.text.x = element_blank(),
          strip.text.x = element_text(size = 12),
          legend.text = element_text(size=10))
#          plot.title = element_text(size=18, face="bold", vjust=2)) + 
#    ggtitle("Water Samples")

legend1 <- get_legend(p1)
p1 <- p1 + theme(legend.position="none")
p2 <- p2 + theme(legend.position="none")
legend2 <- get_legend(p3)
p3 <- p3 + theme(legend.position="none")
p4 <- p4 + theme(legend.position="none")

grid.arrange(p1, p2, legend1, p3, p4, legend2, ncol=3, 
             widths=c(5.2, 5.2, 1.6))

pdf("figS-Richness_by_clade-continent.pdf", width=8, height=5)
grid.arrange(p1, p2, legend1, p3, p4, legend2, ncol=3, 
             widths=c(5.2, 5.2, 1.6))
dev.off()
```



#Alpha Diversity Correlation with Temperature/Latitude
## Tests of correlation
### Temperature with water diversity:

```{r temp_correlation_water}
WvT <- cor.test(adivW$Chao1, adivW$Temperature)
WvT
```

### Temperature with water diversity:

```{r temp_correlation_copepod}
CvT <- cor.test(adivC$Chao1, adivC$Temperature)
CvT
```

### Latitude with water diversity (Spearman's rho)

```{r lat_correlation_water}
WvL <- cor.test(adivW$Chao1, adivW$Latitude, method = "spearman")
WvL
```

### Temperature with water diversity:

```{r lat_correlation_copepod}
CvL <- cor.test(adivC$Chao1, adivC$Latitude, method = "spearman")
CvL
```

```{r corplot_sup}

# Plot correlation between temp/latitude and diversity
a <- melt(adiv, id.vars=c("Location", "Environment"), 
          measure.vars=c("Chao1", "Shannon", "Temperature"))
l <- paste("P = ", signif(CvT$p.value, 2), sep = '')
b <- dcast(a, Location ~ variable + Environment)
div_corplot <- ggplot(b, aes(Chao1_Copepod, Temperature_Copepod)) + 
    geom_point(size=1) + xlab("Copepod Microbiome Diversity") +
    ylab("Water Temperature") + theme(axis.title.y = element_text(angle=90)) +
    geom_smooth(method=lm, fill = "lavenderblush3") + 
    annotate("text", label = l, x = 475, y = 17, size = 4) +
    theme(axis.title.y = element_text(size=14),
          axis.title.x = element_text(size=14),
          plot.title = element_blank())
p1 <- div_corplot

l <- paste("P = ", signif(WvT$p.value, 2), sep = '')
div_corplot <- ggplot(b, aes(Chao1_Water, Temperature_Water)) + 
    geom_point(size=1) + xlab("Water Microbiome Diversity") +
    ylab("Water Temperature") + theme(axis.title.y = element_text(angle=90)) +
    geom_smooth(method=lm, fill = "lavenderblush3") + 
    annotate("text", label = l, x = 3250, y = 17, size = 4) +
    theme(axis.title.y = element_text(size=14),
          axis.title.x = element_text(size=14),
          plot.title = element_blank())
p2 <- div_corplot

a <- melt(adiv, id.vars=c("Location", "Environment"), 
          measure.vars=c("Simpson", "Chao1", "Latitude"))
b <- dcast(a, Location ~ variable + Environment)

l <- paste("P = ", signif(CvL$p.value, 2), sep = '')
div_corplot <- ggplot(b, aes(Chao1_Copepod, Latitude_Copepod)) + 
    geom_point(size=1) + xlab("Copepod Microbiome Diversity") +
    ylab("Latitude") + theme(axis.title.y = element_text(angle=90)) +
    geom_smooth(method=lm, fill = "lavenderblush3") + 
    annotate("text", label = l, x = 475, y = 42, size = 4) +
    theme(axis.title.y = element_text(size=14),
          axis.title.x = element_text(size=14),
          plot.title = element_blank())
p3 <- div_corplot

l <- paste("P = ", signif(WvL$p.value, 2), sep = '')
div_corplot <- ggplot(b, aes(Chao1_Water, Latitude_Water)) + 
    geom_point(size=1) + xlab("Water Microbiome Diversity") +
    ylab("Latitude") + theme(axis.title.y = element_text(angle=90)) +
    geom_smooth(method=lm, fill = "lavenderblush3") + 
    annotate("text", label = l, x = 3250, y = 47, size = 4) +
    theme(axis.title.y = element_text(size=14),
          axis.title.x = element_text(size=14),
          plot.title = element_blank())
p4 <- div_corplot


pdf("figS-Richness_correlation_plots.pdf", height=5.3, width=8)
grid.arrange(p1, p2, p3, p4, ncol=2)
dev.off()

grid.arrange(p1, p2, p3, p4, ncol=2)
```