```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs_betaDiv_relative/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r read_data}
library(DESeq2)
library(phyloseq)
library(RColorBrewer)
library(ggplot2)
library(tidyr)
library(ape)
library(dplyr)
library(reshape2)
library(knitr)

set.seed(9373399)

# Import data
otuTable <- read.csv("data/otus.csv", header = TRUE, row.names = 1)
taxTable <- read.csv("data/taxa.csv", quote = '"')
sampleData <- read.csv("data/sampleData.csv")
tree <- read.tree("data/tree.tre")
rownames(sampleData) <- sampleData[, 1]
sampleData <- sampleData[, -1]
taxmat <- as.matrix(taxTable)
rownames(taxmat) <- taxmat[, 1]
taxmat <- taxmat[, -1]

otuTable <- otu_table(otuTable, taxa_are_rows = FALSE)
taxTable <- tax_table(taxmat)
physeq <- phyloseq(otuTable, taxTable, tree)
sample_data(physeq) <- sampleData

## Variance Stabilized sample counts with DESeq2 (per "waste not, want not" paper)
pseq_deseq <- phyloseq_to_deseq2(physeq, ~ Environment)

# calculate geometric means prior to estimate size factors
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
geoMeans <- apply(counts(pseq_deseq), 1, gm_mean)

#Dispersion Estimates in DESeq2
pseq_vst <- estimateSizeFactors(pseq_deseq, geoMeans = geoMeans)
pseq_vst <- estimateDispersions(pseq_vst)
pseq_vst <- getVarianceStabilizedData(pseq_vst)
pseq_vst[pseq_vst < 0] <- 0

otu_table(physeq) <- otu_table(pseq_vst, taxa_are_rows = TRUE)

rm(pseq_vst, geoMeans, pseq_deseq, sampleData, taxmat, otuTable, taxTable, tree)
```

# Distance testing (perMANOVA (adonis))
# Unifrac Unweighted on all samples (copepods and water)

I want to test the clustering of the points based on treatment. This can be done using the implementation of perMANOVA in the 'adonis' function from the vegan package

### Is the Environment (Host vs. Water) a significant factor?

```{r adonis_env}
library("vegan")
packageVersion("vegan")

phywat_cop <- physeq
unifrac_unweighted <- UniFrac(phywat_cop, weighted=FALSE) 
df <- data.frame(sample_data(phywat_cop))

adonis(unifrac_unweighted ~ Environment, data=df)
```


### Is the salinty (saline vs. fresh) a factor for ALL samples (copepod and water)?

```{r adonis_all_sal}
adonis(unifrac_unweighted ~ WaterType, data=df)
```

---

# Copepod only samples
##Unifrac unweighted testing

```{r adonis_cop}
phywat_cop <- subset_samples(physeq, Environment != "Water")
unifrac_unweighted <- UniFrac(phywat_cop, weighted=FALSE)
df <- data.frame(sample_data(phywat_cop))
```

### Is salinity (Salt vs. Fresh) as significant factor?

```{r adonis_cop_salt}
sal <- adonis(unifrac_unweighted ~ WaterType, data=df)
sal
```

### Is Clade an important factor?

```{r adonis_cop_clade}
clade <- adonis(unifrac_unweighted ~ Clade, data=df)
clade
```

### Is latitude an important factor?

```{r adonis_cop_lat}
lat <- adonis(unifrac_unweighted ~ Latitude, data=df)
lat
```

### Is Temperature an important factor?

```{r adonis_cop_temp}
phywat_cop <- subset_samples(phywat_cop, Location != "Montmagnay")
unifrac_unweighted <- UniFrac(phywat_cop, weighted=FALSE)
df <- data.frame(sample_data(phywat_cop))
temp <- adonis(unifrac_unweighted ~ Temperature, data=df)
temp
```


Based on the previous tests, it seems obvious that there is a salinity (fresh v. salt) effect. Other environmental factors are not so clear. It seems that there could be an effect of latitude on community composition. However there are a lot of confounding factors, including the fact that clade covaries with latitude. I'm also doing a lot of multiple testing. I'll adjust these p-values to the multiple tests that I performed:

```{r adonis_multest}
pvals <- c(sal$aov.tab$`Pr(>F)`[1], clade$aov.tab$`Pr(>F)`[1],
           lat$aov.tab$`Pr(>F)`[1], temp$aov.tab$`Pr(>F)`[1])
p.adjust(pvals, method = "holm")
```

Which is to say that salinity [1], remains a factor after correcting for multiple inferences, but none of the other factors (clade, latitude, temperature; [2:4]) do.

# Water only samples
##Unifrac unweighted testing

```{r adonis_wat}
phywat_cop <- subset_samples(physeq, Environment != "Copepod")
unifrac_unweighted <- UniFrac(phywat_cop, weighted=FALSE)
df <- data.frame(sample_data(phywat_cop))
```

### Is salinity (Salt vs. Fresh) as significant factor?

```{r adonis_wat_salt}
adonis(unifrac_unweighted ~ WaterType, data=df)
```