#E affinis 16S analysis
### Martin Bontrager

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs_foldchange_genus/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r read_data}
library(phyloseq)
library(RColorBrewer)
library(ggplot2)
library(plyr)
library(tidyr)
library(ape)
library(dplyr)
library(reshape2)
library(knitr)

set.seed(9373399)
output_table = FALSE #Set to true to output a csv table of relative genus abundance

otuTable <- read.csv("data/otus.csv", header = TRUE, row.names = 1)
taxTable <- read.csv("data/taxa.csv", quote = '"')
sampleData <- read.csv("data/sampleData.csv")
tree <- read.tree("data/tree.tre")
rownames(sampleData) <- sampleData[, 1]
sampleData <- sampleData[, -1]
taxmat <- as.matrix(taxTable)
rownames(taxmat) <- taxmat[, 1]
taxmat <- taxmat[, -1]


otuTable <- otu_table(otuTable, taxa_are_rows = FALSE)
taxTable <- tax_table(taxmat)

physeq <- phyloseq(otuTable, taxTable, tree)
sample_data(physeq) <- sampleData

rm(taxmat, sampleData, otuTable, taxTable, tree)

# I agglomerate by genus
physeq_1 <- tax_glom(physeq, "Genus")

if (output_table){
    physeq1 <- transform_sample_counts(physeq_1, function(x) round(100 * (x/sum(x)), 2))
    b <- psmelt(physeq1)[, 1:3]
    c <- dcast(b, OTU ~ Sample)
    e <- as.data.frame(tax_table(physeq_1))
    e <- cbind(OTU = rownames(e), e)
    f <- merge(e, c, by = "OTU")
    d <- sapply(f[,1], function(x) gsub("OTU_", "", x), USE.NAMES = FALSE)
    d <- as.numeric(d)
    f$OTU <- sprintf("OTU_%05d", d)
    f <- f[order(f$OTU),]
    write.csv(f, "output/sample_matrix_genus.csv")
    rm(physeq1, b, c, d, e, f)
}

## ggplot theming
theme_set(theme_bw())
pal = "Set1"
scale_colour_discrete <- function(palname = pal, ...) {
        scale_colour_brewer(palette = palname, ...)
}

scale_fill_discrete <- function(palname = pal, ...) {
        scale_fill_brewer(palette = palname, ...)
}

getPalette <- colorRampPalette(brewer.pal(9, "Set1"))

# Plot title size, etc.
fontsize = 18L
theme_update(axis.title.x = element_text(size=fontsize))
theme_update(axis.title.y = element_text(size=fontsize))
theme_update(plot.title = element_text(size=fontsize+2))
```

# Validation
##Multiple testing

I want to know which taxa are responsible for the major differences between copepods and water. I will do that with validation and FDR correction. Taxa that are significantly different between water/copepods samples:

```{r mul}
# Following a procedure online, convert phyloseq to DESeq object
# Perform analyses on "Environment", i.e. which taxa separate copepods and water
library(DESeq2)
packageVersion("DESeq2")
copdds <- phyloseq_to_deseq2(physeq_1, ~ Environment)

# calculate geometric means prior to estimate size factors
gm_mean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}
# calculate geometric means prior to estimate size factors
geoMeans <- apply(counts(copdds), 1, gm_mean)
copdds <- estimateSizeFactors(copdds, geoMeans = geoMeans)
copdds <- DESeq(copdds, fitType="local")

#Output and format results
res <- results(copdds)
res <- res[order(res$padj, na.last=NA), ]
alpha <- 0.01
sigtab <- res[(res$padj < alpha), ]
sigtab <- cbind(as(sigtab, "data.frame"), as(tax_table(physeq_1)[rownames(sigtab), ], "matrix"))

sigtab <- sigtab[,1:12]

# Positive log fold changes in this case mean abundant taxa in water:
posigtab <- sigtab[sigtab[, "log2FoldChange"] > 0, ]
# Negative log fold changes mean abundance in copepods
negsigtab <- sigtab[sigtab[, "log2FoldChange"] < 0, ]
print("Taxa that are more abundant in water:")
kable(head(posigtab, 30))
print("Taxa that are more abundant in copepods:")
kable(head(negsigtab, 30))
```

## Plotting the log fold change
```{r mulplot}
theme_set(theme_bw())
colorCount <- length(levels(sigtab$Genus))
sigtabgen <- sigtab
# Phylum order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Phylum = factor(as.character(sigtabgen$Phylum), levels=names(x))
# Family order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Family = factor(as.character(sigtabgen$Genus), levels=names(x))
ggplot(sigtabgen, aes(x=Genus, y=log2FoldChange, color=Phylum)) + 
    geom_point(size=6) + 
    scale_colour_manual(values = sample(getPalette(colorCount))) +
    theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))
```

## Saltwater vs. Freshwater (Copepods only)

```{r cop_salt_fresh}
# Perform analyses on copepods in saltwater vs freshwater
# , i.e. which taxa are distinct between invasions
# Subset data for plotting purposes
copepods_only <- subset_samples(physeq_1, Environment=="Copepod")

copdds <- phyloseq_to_deseq2(copepods_only, ~ WaterType)
geoMeans <- apply(counts(copdds), 1, gm_mean)
copdds <- estimateSizeFactors(copdds, geoMeans = geoMeans)
copdds <- DESeq(copdds, fitType = "local")

#Output and format results
res <- results(copdds)
res <- res[order(res$padj, na.last=NA), ]
alpha <- 0.05
sigtab <- res[(res$padj < alpha), ]
sigtab <- cbind(as(sigtab, "data.frame"), 
                as(tax_table(copepods_only)[rownames(sigtab), ], "matrix"))
sigtab <- sigtab[,1:12]

posigtab <- sigtab[sigtab[, "log2FoldChange"] > 0, ]
negsigtab <- sigtab[sigtab[, "log2FoldChange"] < 0, ]
kable(head(posigtab, 30))
kable(head(negsigtab, 30))

```

## Plotting the log fold change
```{r mulplot_cop}
theme_set(theme_bw())
colorCount <- length(levels(sigtab$Phylum))
sigtabgen <- sigtab
# Phylum order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$Phylum, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Phylum = factor(as.character(sigtabgen$Phylum), levels=names(x))
# Genus order
x = tapply(sigtabgen$log2FoldChange, sigtabgen$Genus, function(x) max(x))
x = sort(x, TRUE)
sigtabgen$Genus = factor(as.character(sigtabgen$Genus), levels=names(x))
ggplot(sigtabgen, aes(x=Genus, y=log2FoldChange, color=Phylum)) + 
    geom_point(size=6) + 
    scale_colour_manual(values = sample(getPalette(colorCount))) +
    theme(axis.text.x = element_text(angle = -90, hjust = 0, vjust=0.5))
```


## Red Clade across salinity

There are no taxa that differ significantly between water types in Red clade.



## Green Clade across salinity

none

## purple clade across salinity

```{r purpleclade}
# Perform analyses on copepods in saltwater vs freshwater
# , i.e. which taxa are distinct between invasions
# Subset data for plotting purposes
copepods_only <- subset_samples(physeq_1, Environment=="Copepod")
copepods_only <- subset_samples(copepods_only, Region=="Europe")

copdds <- phyloseq_to_deseq2(copepods_only, ~ WaterType)
geoMeans <- apply(counts(copdds), 1, gm_mean)
copdds <- estimateSizeFactors(copdds, geoMeans = geoMeans)
copdds <- DESeq(copdds, fitType = "local")

#Output and format results
res <- results(copdds)
res <- res[order(res$padj, na.last=NA), ]
alpha <- 0.01
sigtab <- res[(res$padj < alpha), ]
sigtab <- cbind(as(sigtab, "data.frame"), 
                as(tax_table(copepods_only)[rownames(sigtab), ], "matrix"))
sigtab <- sigtab[,1:11]

posigtab <- sigtab[sigtab[, "log2FoldChange"] > 0, ]
negsigtab <- sigtab[sigtab[, "log2FoldChange"] < 0, ]
kable(head(posigtab, 30))
kable(head(negsigtab, 30))

```
